\chapter{Suggestions for Voters}

% We want this to fit on one page!
\vspace{-10pt}

The following are several suggestions to think about before voting on
an MPI Forum proposal.  

\begin{enumerate}
\item Actually read the proposal.  Take time to think about it.
  Socialize it with your colleagues.

\item Is this proposal just ``syntactic sugar'' for something that
  could be implemented outside of an MPI implementation?

\item Does this proposal represent a ``best practice''?

\item Is this proposal useful on a wide variety of platforms /
  architectures, both today and in the conceivable future?

\item Are there applications that will use the functionality from this
  proposal?

\item Is the use case for this proposal ephemeral?

\item Implementation issues:
  \begin{enumerate}
  \item Is there an implementation?  Implementations may not be
    {\em required}, but should be highly valued, especially for
    ``non-trivial'' proposals.
    
  \item Does the implementation show a performance or functionality
    benefit that could not be implemented outside of an MPI
    implementation?
    
  \item Is the proposal implementable on a wide variety of platforms /
    architectures?
  \end{enumerate}

\item Proposal quality issues:
  \begin{enumerate}
  \item Is the proposal well-written?
    
  \item Is the proposal too young?  E.g., does this proposal represent
    new work that may not yet have been completely vetted, thought
    through, or simply had time to mature?
    
  \item Is the proposal too old?  E.g., has there been a significant
    time lapse between its reading and/or ballots?  (if so, why?)
  \end{enumerate}

\end{enumerate}
